---
title: "Lab7"
author: "Can Zhang"
date: "2024-10-17"
output: "github_document"
---
# Load packages and dataset
```{r}
library(haven)
library(bruceR)
library(dplyr)
library(ggstatsplot)
library(performance)
library(sjPlot)

ex <- read_sav("C:/Users/User/Documents/GitHub/Lab7/ex.sav")

lab7<-read.csv("C:/Users/User/Documents/GitHub/Lab7/lab7.csv")



```

# Correlation
```{r}
#First, you will need to select variables or composites of interest

ex <- ex %>%
  select(Conscientiousness, AverageHoursofSleep, GPA)

#The Corr function also gives you a correlation plot between all variables, but it doesn't look pretty enough and not very customizable

Corr(ex)

#If you hate the correlation plot design, you can use the following function
ggcorrmat(ex)

#If you want to visualize the correlation between two specifcic variables, you can use the following code
ggplot(ex, aes(x = Conscientiousness, y = GPA)) + geom_point() + geom_smooth() + theme_bruce()

#Note that the line will naturally follow the data points, but you can specify to plot the fittest line by adding method = lm to the geom_smooth() function
ggplot(ex, aes(x = Conscientiousness, y = GPA)) + geom_point() + geom_smooth(method = lm) + theme_bruce()




```

# Regression
```{r}
#step 1: build a model
model<-lm(GPA ~ Conscientiousness + AverageHoursofSleep, data = ex)


#step 2: check the assumptions
check_model(model)

#step 3: summarize results; add ,std = TRUE if you want to get standardized coefficients
model_summary(model)

# or using the tab_model functionl; add  ,show.std = TRUE if you want to get standardized coefficients
tab_model(model, show.std = TRUE)

#step 4: plot the results; change type ="est" to type = "std" if you want to plot the standardized coefficients
plot_model(model,  type ="est",  show.values = TRUE, vline.color = "#1B191999", line.size = 1.5, dot.size = 2.5, colors = "blue") + theme_bruce()

```



# Q1: Use the Corr or the ggcorrmat function to plot the correlation between all variables. Which personality traits are correlated with satisfaction with life and which one is not? 
```{r}
ggcorrmat(lab7)
```
All of the Big 5 personality traits except openness are significantly correlated with SWL, as indicated by the plot.

# Q2: Run a multiple regression using the Big 5 personality traits to predict satisfaction with life and plot the results
```{r}

#for this lab assignment you would assume assumptions are met, but for your own analysis, you will need to examine assumptions carefully
#step 1: build a model
model_lab7<-lm(SWL ~ Extraversion + Agreeableness + Conscientiousness + EmotionalStability + Openness, data = lab7)

# or using the tab_model functionl; add  ,show.std = TRUE if you want to get standardized coefficients
tab_model(model_lab7, show.std = TRUE)

#step 4: plot the results; change type ="est" to type = "std" if you want to plot the standardized coefficients
plot_model(model_lab7,  type ="std",  show.values = TRUE, vline.color = "#1B191999", line.size = 1.5, dot.size = 2.5, colors = "purple") + theme_bruce()

```

# Q3: Interpret the R2. Which personality trait(s) can explain unique variance in satisfaction with life when controlling for each other, and which ones cannot? If someone asks you for advice on how to improve life satisfaction, based on your results, which personality trait would you recommend them to change and why?
Extraversion, emotional stability and openness can explain unique variance in SWL when controlling for each other, and agreeableness and conscientiousness cannot. 
Based on these results, focusing on improving extraversion and emotional stability would likely have the most positive impact on life satisfaction. Since they have strong positive associations with life satisfaction. On contrasts, Openness shows a negative relationship with life satisfaction in this model. This suggests that higher openness might be associated with lower SWL. Thus, improving emotional stability and extraversion rather than increasing openness might be more beneficial for improving life satisfaction.

